### 2020-07-15 manuscript/v1_medrxiv/manuscript.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{medRxiv} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Causal analysis of COVID-19 observational data in German districts reveals effects of mobility, awareness, and temperature},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage{geometry}
% \usepackage[numbers,sort&compress]{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Causal analysis of COVID-19 observational data in German districts
reveals effects of mobility, awareness, and temperature}
    \author[Zi]{Edgar Steiger}
   \ead{esteiger@zi.de} 
    \author[Zi]{Tobias Mußgnug}
   \ead{tmussgnug@zi.de} 
    \author[Zi]{Lars Eric Kroll}
   \ead{lkroll@zi.de} 
      \address[Zi]{Central Research Institute of Ambulatory Health Care in Germany (Zi),
Salzufer 8, D-10587 Berlin, Germany}

  \begin{abstract}
  Mobility, awareness, and weather are suspected to be causal drivers for
  new cases of COVID-19 infection. Correcting for possible confounders, we
  estimated their causal effects on reported case numbers. To this end, we
  used a directed acyclic graph (DAG) as a graphical representation of the
  hypothesized causal effects of the aforementioned determinants on new
  reported cases of COVID-19. Based on this, we computed valid adjustment
  sets of the possible confounding factors. We collected data for Germany
  from publicly available sources (e.g.~Robert Koch Institute, Germany's
  National Meteorological Service, Google) for 401 German districts over
  the period of 15 February to 8 July 2020, and estimated total causal
  effects based on our DAG analysis by negative binomial regression. Our
  analysis revealed favorable causal effects of increasing temperature and
  of increased public mobility for essential shopping (grocery and
  pharmacy) as well as awareness measured by COVID-19 burden, all of them
  reducing the outcome of newly reported COVID-19 cases. Conversely, we
  saw adverse effects of public mobility in retail and recreational areas,
  awareness measured by searches for ``corona'' in Google and higher
  rainfall, leading to an increase in new COVID-19 cases. This
  comprehensive causal analysis of a variety of determinants affecting
  COVID-19 progression gives strong evidence for the driving forces of
  mobility, public awareness, and temperature, whose implications need to
  be taken into account for future decisions regarding pandemic
  management.
  \end{abstract}

 \end{frontmatter}

\section{Introduction}\label{introduction}

As the COVID-19 pandemic progresses, research on mechanisms behind the
transmission of SARS-CoV-2 shows conflicting evidence
\citep{who2020report, Chinazzi395, guan2020clinical}. While effects of
mobility have been extensively discussed, less is known on other factors
such as changing awareness in the population
\citep{higgins_correlations_2020, li_retrospective_2020, yuan_trends_2020}
or the effects of temperature
\citep{bannister-tyrrell_preliminary_2020, demongeot_temperature_2020, liu_impact_2020}.
A limiting factor in many studies is the lack of a causal approach to
assess the causal contributions of various factors
\citep{Greenland1999}. This can lead to distorted estimates of the
causal factors with observational data
\citep{Greenland1999, schipf_directed_2011, textor_robust_2017}.

With COVID-19, we find ourselves in a situation in which information on
the causal contribution of various influencing factors in the population
is urgently needed to inform politicians and health authories. On the
other hand, trials cannot be carried out for obvious ethical and legal
reasons. Therefore, when assessing the effects of determinants of
SARS-CoV-2 spread, special attention must be paid to strategies for the
selection of confounding factors.

Another problem with assessing the effects of various determinants of
SARS-CoV-2 spread is the heterogeneity of the countries and regions
examined for example in the Johns Hopkins University (JHU) COVID-19
database \citep{jhucovid19db2020}. The comparison of time series of case
numbers from different countries and observational periods can be
strongly distorted by different factors like testing capacities and
regional variations.

Our objective is to provide estimates of the causal effects of the main
drivers of the pandemic with reduced bias. We conducted a scoping review
of the available studies regarding signaling pathways and determinants
of the spread of SARS-CoV-2 infections and the reported new COVID-19
cases. Then we integrated the current findings into a directed acyclic
graph for the progress of the pandemic at the regional level. Using the
resulting model and the do-calculus we found identifiable effects
without blocked causal paths whose effects can be analyzed with
observational data \citep{Pearl_2014}. We used regional time series data
of all German districts (\(401\)) from various publicly available
sources to analyze these questions on a regional level. Germany is a
good choice in this regard, because it has ample data on contributing
factors on the regional level and has had high testing and treatment
capacities from early on in the pandemic.

\section{Causal Model}\label{causal-model}

We used a directed acyclic graph (DAG)
\citep{schipf_directed_2011, textor_robust_2017} as a tool to analyze
the causal relationships between several exposures and SARS-CoV-2
spread. To get an overview on published associations, a scoping review
was conducted from 20th to 22nd of May 2020 within Pubmed and Google
scholar. Restrictions were applied to English and German language and
the publication date in the last one year. The following search terms
were applied to abstracts and title in Pubmed (``COVID-19'' OR
``COVID19'' OR ``Corona'' OR ``Coronavirus'' OR ``SARS-CoV-2'') and
connected separately in each case with the exposure variables
(``mobility'', ``public awareness'', ``awareness'', ``google
trends'',``ambient temperature'', ``temperature''). For ``mobility'', we
analyzed \(n=8\) studies, \(N=103\) were scanned in Pubmed, together
with the first ten pages (100 results) in Google scholar
(``awareness''/``public awareness''/``google trends'' \(n=9\),
\(N=215\); ``temperature''/``ambient temperature'' \(n=16\), \(N=235\)).
We integrated these findings where possible into the construction of our
DAG, which can be seen in Figure \ref{fig:dag-covid-19}.

\begin{figure}
\centering
\includegraphics{figures/f_full_covid19_dag.eps}
\caption{\label{fig:dag-covid-19}DAG of determinants of reported COVID-19
cases on the district level. Unobserved variables are light gray,
variables marked with an asterisk (*) are confounded by
weekday/holiday.}
\end{figure}

A number of studies report a strong association of \textbf{mobility}
restrictions on the number of new COVID-19 cases: Restrictive measures
(e.g. ``stay-at-home'' orders, travel bans, or school closures) are
shown to possibly reduce the COVID-19 incidence
\citep{chang_modeling_2020, Chinazzi395, fowler_effect_2020, kraemer_effect_2020, lasry_timing_2020, linka_outbreak_2020, mazzoli_effects_2020, xiong_data-driven_2020}.
However, some studies point out the combination of various
non-pharmaceutical interventions (NPIs) is decisive to prevent new
infections \citep{juni_impact_2020, lai_effect_2020}.

Google Trends \citep{google_trends} data can be used as a tool to get
insights into public interest (\textbf{awareness}) in the coronavirus
disease. Several recent studies imply a connection of relative search
volumes (RSV) indices and reported new COVID-19 cases
\citep{ayyoubzadeh_predicting_2020, effenberger_association_2020, higgins_correlations_2020, li_retrospective_2020, lin_google_2020, mavragani_tracking_2020, walker_use_2020, yuan_trends_2020, zhou_effects_2020}.
Some search terms e.g. ``COVID-19'' or ``coronavirus'' predated newly
infected cases/total number of cases by roughly 7 to 14 days for
different countries
\citep{effenberger_association_2020, higgins_correlations_2020, li_retrospective_2020, yuan_trends_2020}.
Additionally, we acknowledged that individual risk-aware behavior might
be a reaction to the current COVID-19 burden (measured as reported cases
at the day of exposure).

Mixed evidence is available regarding the effect of
\textbf{temperature}: On the one hand several papers report an
association between increase in temperature and decrease in newly
infected COVID-19 cases
\citep{bannister-tyrrell_preliminary_2020, demongeot_temperature_2020, liu_impact_2020, qi_covid-19_2020, shi_impact_2020, sobral_association_2020, tosepu_correlation_2020, Wang2020temperature, wu_effects_2020}.
On the other hand, also the opposite has been found
\citep{auler_evidence_2020, xie_association_2020}. Some studies found no
association at all
\citep{briz-redon_spatio-temporal_2020, iqbal_nexus_2020, jahangiri_sensitivity_2020, juni_impact_2020, yao_no_2020}.
It should be noted that few studies considered other confounding
variables than meteorological ones (especially age and population
density among others
\citep{briz-redon_spatio-temporal_2020, juni_impact_2020, wu_effects_2020}).
In addition, the transferability of results between different climate
zones is questionable. To avoid possible bias caused by weather
variables other than temperature, we included rain, wind, and humidity
in our model.

When investigating causal determinants of SARS-CoV-2 infections, a
number of confounders have to be considered. Well-known risk factors for
SARS-CoV-2 as well as for other infections are demographic factors such
as age, gender, socio-economic status (SES), population density, and
foreign citizenship/ethnicity
\citep{de2020risk, Dragano2020.06.17.20133918, jhucovid19db2020}. In
Germany along with other countries (i.e.~Brazil, USA, or the UK),
populist parties or politicians and their electorate tend to be more
sceptical about effects of containment measures than the other part of
the electorate \citep{dohle_wingen_schreiber_2020, engle_staying_2020}.
Therefore we considered both ``right-wing populist party votes'' and
``voter turnout'' as possible confounders. Public health interventions
were also taken into account (contact restrictions, school closures
etc.), as their implementation showed strong correlations with
controlling the spread of SARS-CoV-2
\citep{cowling2020impact, juni_impact_2020, lai_effect_2020}. To avoid
bias due to reporting delay of case numbers we had to include weekday
and German holidays. We include some unobserved variables in our DAG
(e.g. ``Herd immunity''), too. Please note that ``Exposure to
SARS-CoV-2'' is itself an unobserved variable: German case numbers are
reported with delay after date of exposure and symptom onset.
\emph{Exposure to the virus} should not be confused with the formal
\emph{exposure variables} of the DAG (mobility, awareness, temperature).

\section{Data}\label{data}

We collected and aggregated data on reported COVID-19 cases, regional
socio-demographic factors, weather, and general mobility on district and
state level in Germany for the period of 15 February 2020 to 8 July
2020. Our observation period for the outcome consisted of all dates from
23 February 2020 to 8 July 2020 (\(T=137\)), since we used a lag of 8
days for all confounders. We did not exclude any states or districts
(\(K=401\)). We analyzed the daily reported number of new cases as
outcome (\(K\cdot T=54{\ }937\) observations). The set of possible
predictors was derived from our causal DAG (see Table
\ref{tab:model-vars} and Figure \ref{fig:dag-covid-19}). Due to
modelling and data limitations, some of the predictors were unobserved
or were modelled as a construct consisting of several variables. For our
causal analysis, we computed adjustment sets in three different
scenarios for separate exposures within the DAG: i) mobility of
population, ii) awareness of COVID-19 (i.e.~Google searches for
``corona''), iii) weather (i.e.~temperature).

\subsection{Variables}\label{variables}

We downloaded German daily case numbers on district level reported by
Robert Koch Institute (RKI, \citep{casenumbers_rki}, acquired on 12 July
2020) and aggregated them by date. The number of daily active cases for
day \(d\) was derived by subtracting the total number of reported cases
on day \(d\) and day \(d-14\) (14 days as a conservative estimate for
the infectious period, which corresponds here to the required quarantine
time in Germany).

To assess the mobility of the German population, we used data publicly
available on German state level from Google \citep{google_mobility}.
Measurements are daily relative changes of mobility in percent compared
to the period of 3 January 2020 to 6 February 2020. Missing values
(\(25\) out of \(13{\ }488\)) were imputed with value \(0\) and the
state level measurements were passed onto districts within the
corresponding state. Google mobility data was available for six
different sectors of daily life (``retail and recreation'', ``grocery
and pharmacy'', ``parks'', ``transit stations'', ``workplaces'',
``residential'') which means that ``mobility'' is a construct consisting
of several variables. All variables but ``residential'' mobility are
relative changes of daily visitor numbers to the corresponding sectors
compared to the reference period. ``Residential'' mobility is the
relative change of daily time spent at residential areas.

The notion of awareness in the population of COVID-19 describes the
general state of alertness about the new infectious disease. As such, it
was hard to measure directly. As a proxy, we used the relative interest
in the topic term ``corona'' as indicated by Google searches. The daily
data was available on state level \citep{google_trends} and passed onto
district level. As a second proxy for awareness, we used the daily
reported number of COVID-19 cases on the day of the exposure: Since
media reported case numbers prominently, we assumed that this could
reflect individual awareness, too.

We constructed daily weather from four variables (``temperature'',
``rainfall'', ``humidity'', ``wind''). Weather data was downloaded from
Deutscher Wetterdienst (DWD, \citep{dwd_weather}) for all weather
stations in Germany below 1000 meters altitude with daily records in
2020 until 8 July 2020. District level daily weather data was aggregated
per district by averaging the data from the three nearest weather
stations (which includes weather stations inside the district). Missing
values were imputed with mean values (\(n=59\) for wind).

The reported number of COVID-19 cases varied strongly by day of the
week. Thus, we included ``weekday'' as a categorical variable.
Similarly, the reported cases and the exposure to the virus were
affected by official holidays. Within the observation period, this
included among others Good Friday, Easter Monday, and Labor Day. To
correct for effects of these days, we included two variables in the
model, ``Holiday (report)'' (indicates if the day of the report was a
holiday, because governmental health departments were less likely to be
on full duty) and ``Holiday (exposure)'' (indicates if the day of
exposure to the virus was a holiday, because the population behaves
differently on holidays).

For different official and political measures we used one-hot encoded
daily variables, i.e.~ban of mass gatherings, school and kindergarten
closures and their gradual reopening, contact restrictions, and
mandatory face masks for shopping and public transport.

We included several social, economic, and demographic factors on the
district level with direct or indirect influence on the risk of exposure
to SARS-CoV-2 in our analysis. All are readily available from INKAR
database \citep{inkar}. We used the share of population that is 65 years
or older and the share of population that is younger than 18 years
(Age), the share of females in population (Gender), the population
density, the share of foreign citizenships and the share of the
population seeking refuge (Foreign citizenship), the share of low-income
households (Socio-economic status), voter turnout, share of right-wing
populist party votes, and the number of nursing (retirement) homes.

\begin{table}

\caption{\label{tab:model-vars}Observed model variables}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{llll>{\raggedright\arraybackslash}p{10em}l}
\toprule
variable & dynamics & level & type & unit/comment & source\\
\midrule
Weekday & daily & national & categorical & Sat through Thu as six binary variables, Fri as baseline & -\\
Holiday (report) & daily & national & binary & - & -\\
Holiday (exposure) & daily & national & binary & - & -\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Mobility}}\\
\hspace{1em}retail and recreation & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\hspace{1em}grocery and pharmacy & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\hspace{1em}parks & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\hspace{1em}workplaces & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\hspace{1em}residential & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\hspace{1em}transit stations & daily & state & numeric & percent change compared to reference period & Google \citep{google_mobility}\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Weather}}\\
\hspace{1em}Rainfall & daily & district & numeric & mm (l/sqm) & DWD \citep{dwd_weather}\\
\hspace{1em}Temperature & daily & district & numeric & °C & DWD \citep{dwd_weather}\\
\hspace{1em}Humidity & daily & district & numeric & relative humidity (\%) & DWD \citep{dwd_weather}\\
\hspace{1em}Wind & daily & district & numeric & m/s & DWD \citep{dwd_weather}\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Policies}}\\
\hspace{1em}Ban of mass gatherings & daily & national & binary & - & -\\
\hspace{1em}School and kindergarten closures & daily & state & numeric & 0 for no closure, 1 for full closure, 0.5 for partial reopening & -\\
\hspace{1em}Contact restrictions & daily & national & binary & - & -\\
\hspace{1em}Mandatory face masks & daily & district & binary & - & IZA \citep{mitze2020face}\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Socio-demographic}}\\
\hspace{1em}Age & constant & district & numeric & 2 variables: share of population >=65 years \& <18 years & INKAR \citep{inkar}\\
\hspace{1em}Gender & constant & district & numeric & share of female population & INKAR \citep{inkar}\\
\hspace{1em}Population density & constant & district & numeric & population per sqkm & INKAR \citep{inkar}\\
\hspace{1em}Foreign citizens & constant & district & numeric & 2 variables: share of foreign citizens \& of population seeking refuge & INKAR \citep{inkar}\\
\hspace{1em}Socio-economic status & constant & district & numeric & share of households with low income & INKAR \citep{inkar}\\
\hspace{1em}Turnout & constant & district & numeric & voter turnout in last election & INKAR \citep{inkar}\\
\hspace{1em}Right-wing populist party votes & constant & district & numeric & share of votes for AfD in last election & INKAR \citep{inkar}\\
\hspace{1em}Nursing homes & constant & district & numeric & number of nursing (retirement) homes & INKAR \citep{inkar}\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Awareness}}\\
\hspace{1em}Searches corona & daily & state & numeric & percent relative to other states and observation period & Google \citep{google_trends}\\
\hspace{1em}COVID-19 burden & daily & district & numeric & reported cases on day of exposure & RKI \citep{casenumbers_rki}\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Case numbers}}\\
\hspace{1em}Reported new cases of COVID-19 & daily & district & numeric & - & RKI \citep{casenumbers_rki}\\
\hspace{1em}Active cases & daily & district & numeric & active cases on day of report & RKI \citep{casenumbers_rki}\\
\bottomrule
\end{tabular}
\end{table}

All variables but the outcome ``Reported new cases of COVID-19'' and the
offset ``Active cases'' were centered for numerical stability. We did
not scale variables to unit variance to maintain interpretability of
effects on the original scale of variables. Additionally, we lagged the
effect of all variables (but outcome, offset, and the non-dynamic
socio-demographic variables) by 8 days (see Section \ref{results}) which
means that we assumed that their effects on the outcome will be visible
after 8 days.

\section{Methods}\label{methods}

\subsection{Causal analysis with DAG and adjustment
sets}\label{causal-analysis-with-dag-and-adjustment-sets}

We used a directed acyclic graph as a graphical representation of the
hypothesized causal reasoning that leads to exposure to the SARS-CoV-2
virus, onset of COVID-19, and finally reports of COVID-19 cases. Every
node \(v_i\) in the graph is the graphical representation of an observed
or unobserved variable \(x_i\), a directed edge \(e_{ij}\) is an arrow
from node \(v_i\) to \(v_j\) that implies a direct causal relationship
from variable \(x_i\) onto variable \(x_j\). The set of all nodes is
denoted by \(V\), the set of all edges by \(E\), as such, the complete
DAG is the tuple \(G=(V,E)\). The seminal works of Spirtes and Pearl
\citep{spirtes2000causation, pearl2009causality} introduce the theory of
causal analysis, do-calculus, and how to analyze a DAG to estimate the
total or direct causal effect from a variable \(x_i\) onto a variable
\(x_j\). The direct effect is the effect associated with the edge
\(e_{ij}\) only (if it exists), while the total effect takes indirect
effects via other paths from \(v_i\) to \(v_j\) into account, too. Here
we estimated total effects only, since most of our variables were not
hypothesized to have a direct effect on the \emph{reported} number of
new COVID-19 cases. In contrast to prediction tasks, where one would
include all variables available, it is actually ill-advised to use all
available variables to estimate causal effects, due to introducing bias
by adjusting for unnecessary variables within the causal DAG. This is
why we need to identify a valid set of necessary variables (an
adjustment set) to estimate the proper causal effect
\citep{pearl2009causality}. The ``minimal adjustment set''
\citep{greenland1999causal} is a valid adjustment set of variables that
does not contain another valid adjustment set as a subset. However,
identifying a minimal adjustment set might not be enough to reliably
estimate the causal effect. Thus, we identified the ``optimal adjustment
set'' \citep{henckel2019graphical} as the set of variables which is a
valid adjustment set while having the lowest asymptotic variance in the
resulting causal effect estimates.

We analyzed the DAG from Section \ref{causal-model} with the R Software
\citep{rsoftware} and the R packages \texttt{dagitty} (formal
representation of the graph and minimal adjustment sets
\citep{textor_robust_2017}) and \texttt{pcalg} (for finding an optimal
adjustment set \citep{pcalg}). For the defined exposures and the outcome
``Reported new cases of COVID-19'', we computed the minimal and optimal
adjustment sets. Since it was possible that these sets contained
unobserved variables that needed to be left out of the regression model,
we chose the valid set with the highest pseudo-\(R^2\) (see next
section) to estimate the final total causal effect from exposure to
outcome.

\subsection{Regression with negative binomial
model}\label{regression-with-negative-binomial-model}

We can estimate the causal effect from exposure to outcome by regression
\citep{pearl2009causality}. Since the outcome ``Reported new cases of
COVID-19'' is a count variable, one should not employ a linear
regression model with Gaussian errors, but instead we assumed a
log-linear relationship between the expected value of the outcome \(Y\)
(new cases) and regressors \(x\), as well as a Poisson or negative
binomial distribution for \(Y\):

\begin{equation}
\log(\mathbb{E}[Y|x]) = \alpha + \sum_{i\in S}\beta_i\cdot x_{i}, \label{eq:regmodel}
\end{equation}

where \(\alpha\) is the regression intercept, \(S\) is the set of
adjustment variables for the exposure \(i^{\ast}\) including the
exposure variable itself, \(\beta_i\) are the regression coefficients
corresponding to the variables \(x_i\). As such \(\beta_{i^{\ast}}\) is
the total causal effect from exposure variable \(x_{i^{\ast}}\) on the
outcome Y.

The Poisson regression assumes equality of mean and variance. If this is
not the case one observes so-called overdispersion (the variance is
higher than the mean), this indicates one should use regression with a
negative binomial distribution instead to estimate the variance
parameter separately from the mean.

We needed to account for the fact that our outcome is not counted per
time unit (one day) only, but depends on the number of active COVID-19
cases: Holding all other variables fixed, the number of new cases \(Y\)
is a constant proportion of the number of active cases \(A\). This was
modeled by including an offset \(\log(A+1)\) in the regression model
\eqref{eq:regmodel}:

\begin{align}
\log(\mathbb{E}[Y|x]) &= \alpha + \log(A+1) + \sum_{i\in S}\beta_i\cdot x_{i} \nonumber \\
\Leftrightarrow \log(\frac{\mathbb{E}[Y|x]}{A+1}) &= \alpha + \sum\beta_i\cdot x_{i} \label{eq:logratemodel} \\
\Leftrightarrow \frac{\mathbb{E}[Y|x]}{A+1} &= \exp(\alpha)\cdot \prod \exp(\beta_i)^{x_i}. \label{eq:ratechange}
\end{align}

Here we added a pseudocount ``\(+1\)'' to ensure a finite logarithm and
avoid division by \(0\).

One can interpret the model as approximating the log-ratio of new cases
and active cases by a linear combination of the regressor variables
\eqref{eq:logratemodel}. If all variables \(x_i\) are centered in
\eqref{eq:ratechange}, we have for the baseline
\(\forall i\ x_i=0 \Rightarrow \mathbb{E}[Y|x=0]=\exp(\alpha)\cdot (A+1)\).
In other words, the exponentiated intercept is the baseline daily
infection rate (how many people does one infected individual infect in
one day). If we hold all variables \(x_i\) fixed (e.g.~at baseline 0) in
\eqref{eq:ratechange} but now increase the exposure variable
\(x_{i^{\ast}}=0\) by one unit to \(x_{i^{\ast}}+1=0+1\), we have
\(\mathbb{E}[Y|x']=\exp(\alpha)\cdot(A+1)\exp(\beta_{i^{\ast}}^{x_{i^{\ast}}+1})\prod_{i\neq i^{\ast}}\exp(\beta_i)^0=\exp(\alpha)\cdot(A+1)\exp(\beta_{i^{\ast}})\),
which means the exponentiated coefficient \(\beta_{i^{\ast}}\) describes
the rate change of the outcome by one unit increase of the exposure.

In practice, given observations of \(Y\) and \(x\) we estimate the
regression coefficients \(\alpha\) and \(\beta_i\) by \emph{maximum
likelihood} \citep{maxlikelihood}. Our observational measurements are
\(y_{kt}\) and \(x_{ikt}\), where \(k\) indicates the corresponding
district and \(t\) the date of measurement.

When we analyzed different adjustment sets given by analysis of the
causal DAG (i.e.~the minimal and optimal adjustment sets), we first
checked if the set included unobserved variables. If this was the case
for the optimal adjustment set, we discarded the unobserved variables
from the set and checked if it was still a valid adjustment set
(function \texttt{gac} in package \texttt{pcalg}
\citep{perkovi2015complete}). If a minimal adjustment set contained
unobserved variables, we discarded the whole set. We conducted a
log-linear regression (function \texttt{glm} with
\texttt{family=poisson()} for Poisson regression, and \texttt{glm.nb}
from the \texttt{MASS} package for the negative binomial regression
\citep{mass}) for every remaining valid adjustment set as regressors and
calculated a Pseudo-\(R^2\) given by \(1-V_m/V_0\), where \(V_m\) is the
sum of squared prediction errors of the current model and \(V_0\) is the
sum of squared prediction errors of the null model (intercept and offset
only). That is, our Pseudo-\(R^2\) is \(1\) minus the fraction of
variance unexplained. Finally, we decided for the model/adjustment set
with the highest pseudo-\(R^2\). We report the exponentiated estimated
coefficients along with 99 percent confidence intervals of the
estimates.

\section{Results}\label{results}

Descriptive statistics for the included variables are presented in Table
\ref{tab:descriptive-vars}.

\begin{table}

\caption{\label{tab:descriptive-vars}Descriptive Statistics for observed variables}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{ll}
\toprule
Variable & mean (SD)\\
\midrule
n & 54937\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Mobility}}\\
\hspace{1em}retail and recreation & -27.14 (24.59)\\
\hspace{1em}grocery and pharmacy & -4.04 (23.00)\\
\hspace{1em}parks & 46.33 (57.94)\\
\hspace{1em}workplaces & -22.98 (20.56)\\
\hspace{1em}residential & 8.20 (6.54)\\
\hspace{1em}transit stations & -29.83 (21.22)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Weather}}\\
\hspace{1em}Rainfall & 1.85 (3.98)\\
\hspace{1em}Temperature & 10.74 (5.27)\\
\hspace{1em}Humidity & 67.68 (13.10)\\
\hspace{1em}Wind & 3.64 (1.67)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Policies}}\\
\hspace{1em}Ban of mass gatherings & 0.82 (0.38)\\
\hspace{1em}School and kindergarten closures & 0.54 (0.36)\\
\hspace{1em}Contact restrictions & 0.74 (0.44)\\
\hspace{1em}Mandatory face masks & 0.48 (0.50)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Socio-demographic}}\\
\hspace{1em}Age (pop. 65 and older) & 22.09 (2.74)\\
\hspace{1em}Age (pop. younger 18) & 16.17 (1.25)\\
\hspace{1em}Gender & 0.51 (0.01)\\
\hspace{1em}Population density & 533.75 (701.84)\\
\hspace{1em}Foreign citizens & 10.03 (5.14)\\
\hspace{1em}Foreign citizens (refugees) & 1.88 (1.14)\\
\hspace{1em}Socio-economic status & 30.64 (6.02)\\
\hspace{1em}Turnout & 75.08 (3.79)\\
\hspace{1em}Right-wing populist party votes & 13.39 (5.32)\\
\hspace{1em}Nursing homes & 36.11 (30.69)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Awareness}}\\
\hspace{1em}Searches corona & 27.24 (18.31)\\
\hspace{1em}COVID-19 burden & 3.55 (10.37)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Case numbers (Outcome and offset)}}\\
\hspace{1em}Reported new cases COVID-19 & 3.60 (10.39)\\
\hspace{1em}Active cases & 49.83 (121.96)\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{figures/f_timeseries_cases_and_lograte.eps}
\caption{\label{fig:cases-log}Temporal and district level variation of
outcome (log-scale)}
\end{figure}

In the observational period, the number of daily reported COVID-19 cases
increased till the end of March/beginning of April and continually
decreased afterwards till the beginning of June 2020 with a slight
increase and decrease afterwards (Figure \ref{fig:cases-log}A). On the
other hand, the (log-)ratio of reported cases over active cases
decreased steeply till the mid of April and increased steadily
afterwards with a slight decrease close to the end of the observation
period (Figure \ref{fig:cases-log}B). Both figures examplify a
considerable variation among the districts (light blue points are
individual district's data).

In Germany, we observed a rebound in mobility after the initial
political measures, reductions in incident cases were associated with a
diminishing public interest in COVID-19, and temperatures were overall
increasing (cf.~Figure \ref{fig:timeseries}); with correlations between
temporal progression and mobility in retail and recreation \(r_{A,D}=\)
-0.02, awareness (``Searches corona'') \(r_{A,C}=\) -0.29, and
temperature \(r_{A,B}=\) 0.79.

\begin{figure}
\centering
\includegraphics{figures/f_timeseries_exposures.eps}
\caption{\label{fig:timeseries}Temporal variation of outcome and main
determinants}
\end{figure}

\subsection{Main results}\label{main-results}

\begin{table}

\caption{\label{tab:causaleffects}Causal effect estimates with 99 percent confidence bands}
\centering
\begin{tabular}[t]{lrrr}
\toprule
cause & effect estimate & CI (lower) & CI (upper)\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Mobility}}\\
\hspace{1em}retail and recreation & 1.0076 & 1.0056 & 1.0097\\
\hspace{1em}grocery and pharmacy & 0.9951 & 0.9938 & 0.9965\\
\hspace{1em}parks & 0.9998 & 0.9994 & 1.0003\\
\hspace{1em}transit stations & 0.9975 & 0.9946 & 1.0004\\
\hspace{1em}workplaces & 1.0031 & 1.0002 & 1.0059\\
\hspace{1em}residential & 1.0080 & 0.9987 & 1.0174\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Awareness}}\\
\hspace{1em}Searches corona & 1.0101 & 1.0085 & 1.0118\\
\hspace{1em}COVID-19 burden & 0.9942 & 0.9932 & 0.9951\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Weather}}\\
\hspace{1em}Temperature & 0.9918 & 0.9879 & 0.9958\\
\hspace{1em}Rainfall & 1.0108 & 1.0066 & 1.0150\\
\hspace{1em}Humidity & 1.0022 & 1.0006 & 1.0039\\
\hspace{1em}Wind & 0.9893 & 0.9802 & 0.9985\\
\bottomrule
\end{tabular}
\end{table}

We list the results of our causal analysis for the effects of our
variables in Table \ref{tab:causaleffects}. The estimates are
multiplicative rates of increase/decrease for a one unit increase of the
respective variable: Values above 1 lead to an increase, below 1 to a
decrease of the infection rate. To put these estimates into perspective,
Figure \ref{fig:effectsrange} shows the relative causal effect of the
different exposure variables on the number of reported COVID-19 cases on
a range of sensible values of the exposure variables (95 percent
quantiles of data points).

\begin{figure}
\centering
\includegraphics{figures/f_exposure_vs_effect.eps}
\caption{\label{fig:effectsrange}Relative causal effects of exposures}
\end{figure}

Within our framework, we saw strong effects for mobility in
retail/recreational areas and essential shopping (grocery and pharmacy).
In the former, an increase of 1 percent point mobility compared to the
reference period (03 January to 06 February 2020) leads to an increase
of the daily reported case number by about 0.8 percent. Contrarily, a
corresponding increase of 1 percent point for the areas of
grocery/pharmacy leads to a decrease in the reported case number by
approximately 0.5 percent. Mobility on workplaces showed a small effect
of 0.3 increase in case numbers for every 1 percent point increase in
mobility. Other causal effects of mobility were insubstantial and not
consistent in their direction (99 percent confidence intervals of
estimates include 1). Figure \ref{fig:effectsrange} shows the effects of
mobility on a range of possible values. Thus, we expect an increase of
daily cases by approximately 23 percent if mobility in retail/recreation
reaches baseline levels of 0 percent difference to the reference period.
On the other hand, an increase of mobility for grocery/pharmacy by 10
percent points compared to the reference period leads to a reduction of
the infection rate by approximately 7 percent.

``Awareness'' had two opposite effects on the outcome in our DAG.
Awareness measured by Google searches for \emph{corona} had a positive
effect on the number of reported cases. An one percent point increase of
the state's Google searches (relative to other states and the
observation period) leads to an increase of approximately 1 percent. For
example, if a district shows 10 percent points more relative searches
for \emph{corona} then another one, we expect approximately 11 percent
more infections for this district after 8 days. \emph{COVID-19 burden}
(reported number of cases on day of exposure) affected the outcome
negatively, where every additional daily case in the district leads to a
0.6 percent decrease in newly reported case numbers. The last plot in
Figure \ref{fig:effectsrange} visualizes this relationship: For a local
outbreak with 20 daily cases as COVID-19 burden, we estimate as total
causal effect a subsequent reduction of infection rate by 9 percent.

Within our model, we observed a causal effect of temperature and all
other weather variables. Every increase of 1 degree Celsius in
temperature leads to a reduction of the daily reported case numbers by
approximately 0.8 percent. Similarly strong was the effect of rainfall:
One millimeter (=1 liter per square meter) more rainfall leads to an
increase of reported case numbers by approximately 1.1 percent. We
observe small effects for humidity and wind as well (higher humidity
leading to fewer cases, stronger wind leading to less cases). In
perspective (Figure \ref{fig:effectsrange}), with temperature we expect
an increase by approximately 9 percent at a daily average temperature of
\(0^{\circ}C\). For rainfall, we expect on a very rainy day with 10 mm
rainfall a corresponding increase of the infection rate by approximately
2 percent.

In all cases we opted to use the reduced optimal adjustment set over the
minimal adjustment sets because of higher pseudo-\(R^2\) values (mostly
above 0.3), except for mobility, where the minimal adjustment set had a
higher pseudo-\(R^2\). Notably, these sets always include most of our
socio-demographic variables as confounders (cf.~Table \ref{tab:adjsets})
as well as the policy variables (except for COVID-19 burden). We also
decided for the lag of 8 days based on the highest pseudo-\(R^2\) values
compared to other lags on the chosen adjustment sets. Similarly,
negative binomial regression was chosen over Poisson regression, because
the latter showed overdispersion and overall lower pseudo-\(R^2\)
values.

\begin{table}

\caption{\label{tab:adjsets}Final adjustment sets for causal analysis}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{llllllll}
\toprule
\rotatebox{90}{ } & \rotatebox{90}{Mobility} & \rotatebox{90}{COVID-19 burden} & \rotatebox{90}{Searches corona} & \rotatebox{90}{Temperature} & \rotatebox{90}{Rainfall} & \rotatebox{90}{Humidity} & \rotatebox{90}{Wind}\\
\midrule
Pseudo-R2 & 0.332 & 0.151 & 0.298 & 0.298 & 0.286 & 0.306 & 0.286\\
Weekday & x & x & x & x & x & x & x\\
Holiday (report) & x & x & x & x & x & x & x\\
Holiday (exposure) & x & x & x & x & x & x & x\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Mobility}}\\
\hspace{1em}retail and recreation &  &  &  &  &  &  & \\
\hspace{1em}grocery and pharmacy &  &  &  &  &  &  & \\
\hspace{1em}parks &  &  &  &  &  &  & \\
\hspace{1em}workplaces &  &  &  &  &  &  & \\
\hspace{1em}residential &  &  &  &  &  &  & \\
\hspace{1em}transit stations &  &  &  &  &  &  & \\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Weather}}\\
\hspace{1em}Rainfall & x & x & x & x &  & x & x\\
\hspace{1em}Temperature & x & x & x &  &  & x & \\
\hspace{1em}Humidity &  & x & x &  &  &  & \\
\hspace{1em}Wind &  & x & x & x & x & x & \\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Policies}}\\
\hspace{1em}Ban of mass gatherings & x &  & x & x & x & x & x\\
\hspace{1em}School and kindergarten closures & x &  & x & x & x & x & x\\
\hspace{1em}Contact restrictions & x &  & x & x & x & x & x\\
\hspace{1em}Mandatory face masks &  &  & x & x & x & x & x\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Socio-demographic}}\\
\hspace{1em}Age & x & x & x & x & x & x & x\\
\hspace{1em}Gender & x & x & x & x & x & x & x\\
\hspace{1em}Population density & x & x & x & x & x & x & x\\
\hspace{1em}Foreign citizens &  & x & x &  &  &  & \\
\hspace{1em}Socio-economic status & x & x & x & x & x & x & x\\
\hspace{1em}Turnout &  & x &  &  &  &  & \\
\hspace{1em}Right-wing populist party votes &  & x & x & x & x & x & x\\
\hspace{1em}Nursing homes &  & x & x & x & x & x & x\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Awareness}}\\
\hspace{1em}Searches corona & x &  &  & x & x & x & x\\
\hspace{1em}COVID-19 burden & x &  & x & x & x & x & x\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}\label{discussion}

\subsection{Main findings}\label{main-findings}

Our objective was to identify causal effects for COVID-19 cases. We
found that weather affects the reported number of infections, especially
temperature (which has a reducing effect on case numbers) and rainfall
(which increases case numbers). We saw that reports of high case numbers
in districts led to a reduction in new infection numbers, which
indicates risk-averse awareness in the population. The overall effect of
mobility showed no consistent effect, however, in specific areas
significant causal effects could be measured: Increasing activity in
retail and recreational areas increased reported case numbers, while
increased movement for essential shopping (grocery and pharmacy) led to
reduced case numbers.

Furthermore, we made a strong case for the use of causal DAGs in
epidemiology and a pandemic like COVID-19: DAGs allow to choose
confounders for the analysis in a principled and statistically correct
way while reducing possible causes for bias. Also, the DAG formalization
allows for discussion about the underlying causal structure.

\subsection{Comparison with previous
research}\label{comparison-with-previous-research}

Most research on determinants affecting case numbers of COVID-19 is
restricted to single aspects
\citep{fowler_effect_2020, li_retrospective_2020, shi_impact_2020, Wang2020temperature}.
To reliably identify causal drivers, one needs to adjust for
confounders. To this end, we used an integrated model with variables
from different aspects like mobility, awareness, weather, or
socio-demographics and identified confounders by causal analysis with a
directed acyclic graph. A causal approach is used in another current
COVID-19 analysis \citep{gencoglu2020causal}. There, however, they
identify the causal relationships (reconstruct a DAG), while we
estimated causal effects for a given hypothesized DAG.

Several studies assessing the impact of public health measures on
mobility have each observed a downward trend accompanied by a decrease
in the number of newly reported cases
\citep{chang_modeling_2020, cowling2020impact, fowler_effect_2020, kraemer_effect_2020, lai_effect_2020, linka_outbreak_2020}.

Our findings regarding awareness/Google Trends analysis are in good
agreement with the correlations found by
\citet{effenberger_association_2020}, \citet{higgins_correlations_2020},
and \citet{yuan_trends_2020}, who conclude that alertness to COVID-19
rises several days before the highest number of cases are reported. At
this point it should be noted, that awareness is substantially
influenced by public media coverage, which should be considered, if
possible, in future studies \citep{higgins_correlations_2020}. As such,
awareness is difficult to measure and here the number of Google searches
for ``corona'' could only be a proxy for this concept.

In addition, in alignment with other recent published studies, our
results confirm evidence which associated a negative effect of
temperature on new COVID-19 cases
\citep{bannister-tyrrell_preliminary_2020, demongeot_temperature_2020, liu_impact_2020, qi_covid-19_2020, shi_impact_2020, sobral_association_2020, tosepu_correlation_2020, Wang2020temperature, wu_effects_2020}.
It is however controversial to other scientific literature describing no
effects
\citep{briz-redon_spatio-temporal_2020, iqbal_nexus_2020, jahangiri_sensitivity_2020, juni_impact_2020, yao_no_2020}
or even converse correlations
\citep{auler_evidence_2020, xie_association_2020}. The conflicting
results might be explained by different climates and characteristics of
the populations under study. While we are confident that our strict
causal analysis resulted in effect estimates as undistorted as possible,
there might be unconsidered bias in those other studies. Further
research needs to be done to elucidate the biological characteristics of
the novel virus SARS-CoV-2 regarding its ambient temperature survival
and transmission. In regards of humidity, our findings (higher humidity
leading to increased case numbers) are in agreement with previous
research \citep{qi_covid-19_2020, wu_effects_2020, juni_impact_2020}.
Finally, we found a positive causal effect of increment precipitation
and a raise in COVID-19 cases, which supports previous observations
\citep{sobral_association_2020}.

\subsection{Limitations and strengths}\label{limitations-and-strengths}

While use of a causal DAG is itself a strong tool to identify causal
effects (and not just statistical associations), it introduces two
limitations: causal assumptions within the graph (depicted by edges)
need to be well justified, and the statistical regression model that
calculates total causal effects needs to be appropriate for the task at
hand. We endorse our graph as a basis for discussion on residual
confounding.

We observed overdispersion and a substantial increase in model
performance with a negative binomial regression compared to Poisson
regression, which is in line with the results on COVID-19 daily case
counts of \citet{kraemer_effect_2020} and others
\citep{liu_impact_2020, bannister-tyrrell_preliminary_2020}. We did not
model case counts with a differential equation model like the classic
SIR-model \citep{kermack1991contributions} and its successors, since
these are more suited to prediction \citep[e.g.][]{rkimodellierung}
while our choice of a negative binomial regression framework allowed us
to estimate the effects of confounders more reliably. There are more
advanced statistical methods for count data, e.g.~zero-inflated models
and mixed models. We tested both approaches as extensions to the
negative binomial regression and experienced numerical problems and
increased computing time, along with an insubstantial increase in model
performance. Furthermore, our model assumed that all variables have
effects proportional to the size of their measurements. It is possible
that some variables show saturation effects or opposite effects for
lower, medium, or high values. This could be modelled with polynomial or
other transformations of the variables, which we did not employ due to
limited temporal and spatial data availability.

As we were using administrative data for our analysis, the results are
susceptible to the Modifiable Area Unit Problem (MAUP)
\citep{Openshaw1984}. The MAUP postulates that different regional
aggregations of the units of observation may lead to different results
and conclusions. Due to limited available data for the different
variables, there is currently no way to overcome this.

Our observation period was restricted to succession from late winter to
spring and summer (February to July). Nevertheless, this transition with
increasing temperature was a natural experiment that allowed clues on
weather effects.

We could not include data on health care utilization during the pandemic
into our models due to the lack of available resources. This is planned
for a later follow up to this paper since we rank health care
utilization and mobility within health care facilities among the strong
factors for COVID-19 progression: personnel in hospitals and private
practices is particularly exposed to infection, while the lack of
adequate care for other diseases has severe effects on general health of
the population. At the same time, health care facilities are key for
testing and surveillance of COVID-19 patients.

While our analysis focused on Germany and its districts, we assume that
results may be transferred to other countries by adjusting for their
respective weather conditions, mobility habits, socio-demographic
characteristics, and other determining factors.

The code and resources for our analysis are available on Github, we
invite other researchers to replicate our analysis with different
assumptions using the files provided in the repository\footnote{The
  repository is located here: https://github.com/zidatalab/causalcovid19}
of the article.

\subsection{Discussion of causal
effects}\label{discussion-of-causal-effects}

In our analysis, the adverse effect of mobility in retail/recreation and
the favorable effect of mobility in grocery/pharmacy make a strong case
for the successful implementation of policies like contact restrictions
which limit the number of individual interactions that lead to
infections. This is due to retail/recreational areas encompassing mostly
places of social gatherings like restaurants and bars, malls, sports and
music venues, among others, while if people are doing more of their
essential shopping at supermarkets, they will most likely stay at home
with less contact to other people.

The causal effects of awareness measured via searches for ``corona'' and
the COVID-19 burden are harder to interpret. We assume that within our
model, the searches for ``corona'' are an insufficient proxy for
awareness, while the decreasing effect for future case numbers of high
daily COVID-19 burden numbers indicates it affects individual
risk-behavior as a deterrent.

Similarly, the effects of temperature and rainfall can be interpreted as
causal effects for indoor and outdoor activities, such that higher
temperatures and low rainfall indicate more people spending time outdoor
while lower temperatures and high rainfall result in indoor activities,
which lead to more infections. In this light, we advocate for
precautious measures like increased hygiene, face masks, and air
ventilation for unavoidable indoor activities.

\subsection{Conclusions}\label{conclusions}

To the best of our knowledge, this is the most comprehensive analysis of
causes for COVID-19 infections which integrates different data sources
(all publicly available). Causal reasoning with a DAG allows us to
estimate the causal effects reliably.

Our findings suggest that the causal effects of mobility, awareness, and
weather need to be taken strongly into account when deciding for
mitigation and suppression measures, depending on the recent and future
COVID-19 pandemic development.

\renewcommand\refname{References}
\bibliography{mybibfile.bib}

\end{document}