### 2020-07-23 manuscript/v2_medrxiv/results.Rmd
fixes tobias and aerosol reference WHO

# Results

Descriptive statistics for the included variables are presented in Table \@ref(tab:descriptive-vars).

```{r descriptive-vars, echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("kableExtra")
library("tidyverse")
library("tableone")

Modeldata <- read_csv("../../data/Modeldata_raw.csv") %>%
  rename(`Gender`=`Sex`,
         `Foreign citizens`=`Foreign residents`,
         `Foreign citizens (refugees)`=`Foreign residents (refugees)`) %>%
  filter(date<="2020-07-08")

t1_data <- Modeldata %>% dplyr::select(-contains("Weekday"), -contains("Holiday"), -`Relaxation of measures`,
                     -contains("Raumeinheit"),
                     -contains("Aggregat"),
                     -contains("iso_"),
                     -contains("census_"),
                     -contains("mean"),
                     -id,-bl_id,
                     -cases, -deaths,-recovered,-daycount) 
t1_data_tab <- CreateTableOne(data = t1_data) 
mymeans <- colMeans(t1_data[, -1], na.rm = TRUE)
# t1.df <- t1_data_tab.csv %>% as_tibble() %>% mutate(Variable=as.character(rownames(t1_data_tab.csv))) %>% 
#   dplyr::select(Variable,Descriptives=Overall)
t1.df <- print(t1_data_tab, quote = FALSE, noSpaces = TRUE, printToggle = FALSE, explain=FALSE) 
write.csv(t1.df,"tables/t_descriptives.csv")
t1.df <- read.csv("tables/t_descriptives.csv") %>% rename("Variable"="X") %>%
  mutate(Variable=recode(Variable,
                         "Mobility (retail and recreation)"="retail and recreation",
                                                          "Mobility (grocery and pharmacy)"="grocery and pharmacy",
                                                          "Mobility (parks)"="parks",
                                                          "Mobility (transit stations)"="transit stations",
                                                          "Mobility (workplaces)"="workplaces",
                                                          "Mobility (residential)"="residential",
                                                          "Weather (rainfall)"="Rainfall",
                                                          "Weather (wind)"="Wind",
                                                          "Weather (temperature)"="Temperature",
                                                          "Weather (humidity)"="Humidity",
                                                          "School/Kita closures"="School and kindergarten closures")) %>%
  mutate(Variable =  factor(Variable, ordered=TRUE, levels = c("n",
                                                 "retail and recreation",
                                                 "grocery and pharmacy",
                                                 "parks",
                                                 "workplaces",
                                                 "residential",
                                                 "transit stations",
                                                 "Rainfall",
                                                 "Temperature",
                                                 "Humidity",
                                                 "Wind",
                                                 "Ban of mass gatherings",
                                                 "School and kindergarten closures",
                                                 "Contact restrictions",
                                                 "Mandatory face masks",

                                                 "Age (pop. 65 and older)",
                                                 "Age (pop. younger 18)",
                                                 "Gender",
                                                 "Population density",
                                                 "Foreign citizens",
                                                 "Foreign citizens (refugees)",
                                                 "Socio-economic status",
                                                 "Turnout",
                                                 "Right-wing populist party votes",
                                                 "Nursing homes",
                                                 "Searches corona",
                                                 "COVID-19 burden", 
                                                 "Reported new cases COVID-19",
                                                 "Active cases"))) %>%
  arrange(Variable)    
write.csv(t1.df,"tables/t_descriptives.csv")

# number of districts
ndistricts <- Modeldata %>% count(id) %>% count() %>% pull(n)

# Obs per district
nobs_per_district <- Modeldata %>% group_by(id) %>% count() %>% ungroup() %>% summarise(mean=mean(n),sd=sd(n))

# Correlation of time and main vars
mycorrs <- Modeldata %>% dplyr::select(date,contains("cases"),contains("Weather"),`Searches corona`,contains("Mobility"), `COVID-19 burden`) %>% 
  mutate(outcome=`Reported new cases COVID-19`/(`Active cases`+1),date=as.numeric(date)) %>% 
  dplyr::select(-contains("cases")) %>% filter(!is.na(outcome)) %>% cor(.)

# show table of decriptives
knitr::kable(readr::read_csv("tables/t_descriptives.csv") %>% dplyr::select("Variable","Overall") %>% rename(`mean (SD)`=`Overall`),
             format = "latex", booktabs = TRUE,
             caption = "Descriptive Statistics for observed variables") %>% kable_styling(font_size = 7, full_width = FALSE) %>%
  pack_rows("Mobility", 2, 7) %>%
  pack_rows("Weather", 8, 11) %>%
  pack_rows("Policies", 12, 15) %>%
  pack_rows("Socio-demographic", 16, 25) %>%
  pack_rows("Awareness", 26, 27) %>%
  pack_rows("Case numbers (Outcome and offset)", 28, 29)
```

```{r cases-log, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temporal and district level variation of outcome (log-scale)"}
knitr::include_graphics(rep('figures/f_timeseries_cases_and_lograte.eps'))
```

In the observational period, the number of daily reported COVID-19 cases increased till the end of March/beginning of April and continually decreased afterwards till the beginning of June 2020 with a slight increase and decrease afterwards (Figure \@ref(fig:cases-log)A). On the other hand, the (log-)ratio of reported cases over active cases decreased steeply till the mid of April and increased steadily afterwards with a slight decrease close to the end of the observation period (Figure \@ref(fig:cases-log)B). Both figures examplify a considerable variation among the districts (light blue points are individual district's data).

In Germany, we observed a rebound in mobility after the initial political measures, reductions in incident cases were associated with a diminishing public interest in COVID-19, and temperatures were overall increasing (cf. Figure \@ref(fig:timeseries)); with correlations between temporal progression and mobility in retail and recreation $r_{A,D}=$ `r round(mycorrs["date", "Mobility (retail and recreation)"], 2)`, awareness ("Searches corona") $r_{A,C}=$ `r round(mycorrs["date", "Searches corona"], 2)`, and temperature $r_{A,B}=$ `r round(mycorrs["date", "Weather (temperature)"], 2)`.

```{r timeseries, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temporal variation of outcome and main determinants"}
knitr::include_graphics(rep('figures/f_timeseries_exposures.eps'))
```

## Main results

```{r causaleffects, echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("kableExtra")
causaleffects <- bind_rows(readr::read_csv("tables/t_effects_Mobility_MinAdjSet1.csv"),
                           readr::read_csv("tables/t_effects_Searches corona_RedOptAdjSet.csv"),
                           readr::read_csv("tables/t_effects_COVID-19 burden_RedOptAdjSet.csv"),
                           readr::read_csv("tables/t_effects_Temperature_RedOptAdjSet.csv"),
                           readr::read_csv("tables/t_effects_Rainfall_RedOptAdjSet.csv"),
                           readr::read_csv("tables/t_effects_Humidity_RedOptAdjSet.csv"),
                           readr::read_csv("tables/t_effects_Wind_RedOptAdjSet.csv")) %>%
               rename(cause=variables, `effect estimate`=estimates, `CI (lower)`=conflow, `CI (upper)`=confhigh)
## Note: no special chars in caption...
knitr::kable(causaleffects %>% 
               mutate(cause=stringr::str_remove(stringr::str_remove(cause,"Mobility \\("),"\\)")),
             booktabs = TRUE,
             caption = "Causal effect estimates with 99 percent confidence bands",
             digits=4) %>%
  kable_styling() %>%
  pack_rows("Mobility", 1, 6) %>%
  pack_rows("Awareness", 7, 8) %>%
  pack_rows("Weather", 9, 12)
```

We list the results of our causal analysis for the effects of our variables in Table \@ref(tab:causaleffects). The estimates are multiplicative rates of increase/decrease for a one unit increase of the respective variable: Values above 1 lead to an increase, below 1 to a decrease of the infection rate. To put these estimates into perspective, Figure \@ref(fig:effectsrange) shows the relative causal effect of the different exposure variables on the number of reported COVID-19 cases on a range of sensible values of the exposure variables (95 percent quantiles of data points).

```{r effectsrange, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Relative causal effects of exposures"}
knitr::include_graphics(rep('figures/f_exposure_vs_effect.eps'))
```

Within our framework, we saw significant effects for mobility in retail/recreational areas and essential shopping (grocery and pharmacy). For retail/recreation, an increase of 1 percent point mobility compared to the reference period (03 January to 06 February 2020) leads to an increase of the daily reported case number by about `r round(100*abs(1-causaleffects%>%filter(cause=="Mobility (retail and recreation)")%>%pull("effect estimate")), 1)` percent. Contrarily, a corresponding increase of 1 percent point for the areas of grocery/pharmacy leads to a decrease in the reported case number by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Mobility (grocery and pharmacy)")%>%pull("effect estimate")), 1)` percent. Mobility on workplaces showed a small effect of `r round(100*abs(1-causaleffects%>%filter(cause=="Mobility (workplaces)")%>%pull("effect estimate")), 1)` increase in case numbers for every 1 percent point increase in mobility. Other causal effects of mobility were insubstantial and not consistent in their direction (99 percent confidence intervals of estimates include 1). Figure \@ref(fig:effectsrange) shows the effects of mobility on a range of possible values. Thus, we expect an increase of daily cases by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Mobility (retail and recreation)")%>%pull("effect estimate"))^(0-mymeans["Mobility (retail and recreation)"])), 0)` percent if mobility in retail/recreation reaches baseline levels of 0 percent difference to the reference period. On the other hand, an increase of mobility for grocery/pharmacy by 10 percent points compared to the reference period leads to a reduction of the infection rate by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Mobility (grocery and pharmacy)")%>%pull("effect estimate"))^(10-mymeans["Mobility (grocery and pharmacy)"])), 0)` percent.

"Awareness" had two opposite effects on the outcome in our DAG. Awareness measured by Google searches for *corona* had a positive effect on the number of reported cases. An one percent point increase of the state's Google searches (relative to other states and the observation period) leads to an increase of approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate")), 1)` percent. For example, if a district shows 10 percent points more relative searches for *corona* than another one, we expect approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate"))^(10)), 0)` percent more infections for this district after 8 days.  _COVID-19 burden_ (reported number of cases on day of exposure) affected the outcome negatively, where every additional daily case in the district leads to a `r round(100*abs(1-causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate")), 1)` percent decrease in newly reported case numbers. The last plot in Figure \@ref(fig:effectsrange) visualizes this relationship: For a local outbreak with 25 daily cases as COVID-19 burden, we estimate as total causal effect a subsequent reduction of infection rate by `r round(100*abs(1-(causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate"))^(25-mymeans["COVID-19 burden"])), 1)` percent.

Within our model, we observed a causal effect of temperature and all other weather variables. Every increase of 1 degree Celsius in temperature leads to a reduction of the daily reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate")), 1)` percent. On the other hand, we found an increasing effect of rainfall: One millimeter (=1 liter per square meter) more rainfall leads to an increase of reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate")), 1)` percent. We observe effects for humidity and wind as well (higher humidity leading to more cases, stronger wind leading to less cases). In perspective (Figure \@ref(fig:effectsrange)), with temperature we expect an increase by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate"))^(0-mymeans["Weather (temperature)"])), 1)` percent at a daily average temperature of $0^{\circ}C$. For rainfall, we expect on a rainy day with 10 mm rainfall a corresponding increase of the infection rate by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate"))^(10-mymeans["Weather (rainfall)"])), 1)` percent.

In all cases we opted to use the reduced optimal adjustment set over the minimal adjustment sets because of higher pseudo-$R^2$ values (mostly above 0.3), except for mobility, where the minimal adjustment set had a higher pseudo-$R^2$. Notably, these sets always include most of our socio-demographic variables as confounders as well as the policy variables (cf. Table \@ref(tab:adjsets), with the exception being COVID-19 burden). We also decided for the lag of 8 days based on the highest pseudo-$R^2$ values compared to other lags on the chosen adjustment sets. Similarly, negative binomial regression was chosen over Poisson regression, because the latter showed overdispersion and overall lower pseudo-$R^2$ values.

```{r adjsets, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
myvariables <- readr::read_delim("tables/t_variables.csv", delim=";")[c(1:22, 24:28), ] %>% dplyr::select(variable) %>%
  mutate(Mobility=" ", `Searches corona`=" ", `COVID-19 burden`=" ", Temperature=" ", Rainfall=" ", Humidity=" ", Wind=" ")
for (cause in colnames(myvariables)[-c(1,2)]) {
  adjset <- readLines(paste0("tables/t_adjust_", cause, "_RedOptAdjSet.csv"))
  for (as in adjset) {
    myvariables[[cause]][grep(as, myvariables$variable)] <- "x"
  }
}
cause <- colnames(myvariables)[2]
adjset <- readLines(paste0("tables/t_adjust_", cause, "_MinAdjSet1.csv"))
for (as in adjset) {
  myvariables[[cause]][grep(as, myvariables$variable)] <- "x"
}
t_pseudor2s <- read_csv("tables/t_pseudor2s.csv", col_names = c("exposure", "pseudor2"), col_types = "cn") %>%
  mutate(pseudor2=as.character(round(pseudor2, 3))) %>%
  pivot_wider(names_from = "exposure", values_from = "pseudor2")

myvariables <- bind_rows(t_pseudor2s, myvariables)[c(8, 1:7)]
myvariables$variable[1] <- "Pseudo-R2"
kable(myvariables %>% rename(" "=variable),
             format = "latex", booktabs = TRUE, 
             caption = "Final adjustment sets for causal analysis") %>%
  kable_styling(font_size=7) %>% # , latex_options="scale_down"
  row_spec(0, angle = 90) %>%
  pack_rows("Mobility", 5, 10) %>%
  pack_rows("Policies", 15, 18) %>%
  pack_rows("Socio-demographic", 19, 26) %>%
  pack_rows("Awareness", 27, 28) %>%
  pack_rows("Weather", 11, 14) # %>%  add_header_above(c(" " = 2, "Awareness" = 2, "Weather" = 4))
```