### 2021-03-01 manuscript/v4_medrxiv_revision/causal_model.Rmd
add new folders for revision, test latexdiff

# Causal Model

We used a directed acyclic graph (DAG) [@schipf_directed_2011;@textor_robust_2017] as a tool to analyze the causal relationships between several exposures and SARS-CoV-2 spread. To get an overview on published associations, a scoping review was conducted from 20th to 22nd of May 2020 within Pubmed and Google scholar. Restrictions were applied to English and German language and the publication date in the last one year. The following search terms were applied to abstracts and title in Pubmed ("COVID-19" OR "COVID19" OR "Corona" OR "Coronavirus" OR "SARS-CoV-2") and connected separately in each case with the exposure variables ("mobility", “public awareness”, "awareness", “google trends”,“ambient temperature”, "temperature"). For "mobility", we analyzed $n=8$ studies, $N=103$ were scanned in Pubmed, together with the first ten pages (100 results) in Google scholar ("awareness"/"public awareness"/"google trends" $n=9$, $N=215$; "temperature"/"ambient temperature" $n=16$, $N=235$). We integrated these findings where possible into the construction of our DAG, which can be seen in Figure \@ref(fig:dag-covid-19).

```{r dag-covid-19, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="DAG of determinants of reported COVID-19 cases on the district level. Unobserved variables are light gray, variables marked with an asterisk (*) are confounded by weekday/holiday.", fig.height=5}
knitr::include_graphics(rep('figures/f_full_covid19_dag.eps'))
```

A number of studies report a strong association of **mobility** restrictions on the number of new COVID-19 cases: Restrictive measures (e.g. “stay-at-home” orders, travel bans, or school closures) are shown to possibly reduce the COVID-19 incidence [@chang_modeling_2020;@Chinazzi395;@fowler_effect_2020;@kraemer_effect_2020;@lasry_timing_2020;@linka_outbreak_2020;@mazzoli_effects_2020;@xiong_data-driven_2020]. However, some studies point out the combination of various non-pharmaceutical interventions (NPIs) is decisive to prevent new infections [@juni_impact_2020;@lai_effect_2020].

Google Trends [@google_trends] data can be used as a tool to get insights into public interest (**awareness**) in the coronavirus disease. Several recent studies imply a connection of relative search volumes (RSV) indices and reported new COVID-19 cases [@ayyoubzadeh_predicting_2020;@effenberger_association_2020;@higgins_correlations_2020;@li_retrospective_2020;@lin_google_2020;@mavragani_tracking_2020;@walker_use_2020;@yuan_trends_2020;@zhou_effects_2020]. Some search terms e.g. "COVID-19" or "coronavirus" predated newly infected cases/total number of cases by roughly 7 to 14 days for different countries [@effenberger_association_2020; @higgins_correlations_2020;@li_retrospective_2020;@yuan_trends_2020]. Additionally, we acknowledged that individual risk-aware behavior might be a reaction to the current COVID-19 burden (measured as reported cases at the day of exposure).

Mixed evidence is available regarding the effect of **temperature**: On the one hand several papers report an association between increase in temperature and decrease in newly infected COVID-19 cases [@bannister-tyrrell_preliminary_2020;@demongeot_temperature_2020;@liu_impact_2020;@qi_covid-19_2020;@shi_impact_2020;@sobral_association_2020;@tosepu_correlation_2020;@Wang2020temperature;@wu_effects_2020]. On the other hand, also the opposite has been found [@auler_evidence_2020;@xie_association_2020]. Some studies found no association at all [@briz-redon_spatio-temporal_2020;@iqbal_nexus_2020;@jahangiri_sensitivity_2020;@juni_impact_2020;@yao_no_2020]. It should be noted that few studies considered other confounding variables than meteorological ones (especially age and population density among others [@briz-redon_spatio-temporal_2020;@juni_impact_2020;@wu_effects_2020]). In addition, the transferability of results between different climate zones is questionable. To avoid possible bias caused by weather variables other than temperature, we included rain, wind, and humidity in our model.

When investigating causal determinants of SARS-CoV-2 infections, a number of confounders have to be considered. Well-known risk factors for SARS-CoV-2 as well as for other infections are demographic factors such as age, gender, socio-economic status (SES), population density, and foreign citizenship/ethnicity [@de2020risk;@Dragano2020.06.17.20133918;@jhucovid19db2020]. In Germany along with other countries (i.e. Brazil, USA, or the UK), populist parties or politicians and their electorate tend to be more sceptical about effects of containment measures than the other part of the electorate [@dohle_wingen_schreiber_2020;@engle_staying_2020]. Therefore we considered both "right-wing populist party votes" and "voter turnout" as possible confounders. Public health interventions were also taken into account (contact restrictions, school closures etc.), as their implementation showed strong correlations with controlling the spread of SARS-CoV-2 [@cowling2020impact;@juni_impact_2020;@lai_effect_2020]. To avoid bias due to reporting delay of case numbers we had to include weekday and German holidays. We include some unobserved variables in our DAG (e.g. "Herd immunity"), too. Please note that "Exposure to SARS-CoV-2" is itself an unobserved variable: German case numbers are reported with delay after date of exposure and symptom onset. *Exposure to the virus* should not be confused with the formal *exposure variables* of the DAG (mobility, awareness, temperature).